\documentclass{ctexart}

\ctexset{
  section = {
    format = {\Large\bfseries\raggedright}, 
    indent = {0pt}                         
  }
}

\usepackage{enumitem}
\usepackage[a4paper, margin=1in]{geometry}
\usepackage{xcolor} 
\usepackage[cache=true]{minted} 
\usepackage[most]{tcolorbox}
\usepackage{tikz}
\tcbuselibrary{minted, skins}

\setminted{
    fontsize=\small,
    style=default,
    linenos=false,
    breaklines=true,
    fontfamily=tt
}

\newtcolorbox{problembox}[2]{
    enhanced,
    sharp corners,
    boxrule=1.5pt,
    colframe=orange!90!black,
    colback=white,
    colbacktitle=white,
    coltitle=black,
    fonttitle=\bfseries,
    title={Problem (#1): #2},
    titlerule=0pt,
    overlay={
        \draw[black, line width=0.4pt] ([xshift=15pt, yshift=3pt]title.south west) -- ([xshift=-15pt, yshift=3pt]title.south east);
    },
    toptitle=6pt,
    bottomtitle=6pt,
    top=8pt,
    left=10pt,
    right=10pt,
    bottom=10pt
}

\newtcolorbox{tipbox}[1]{
    enhanced,
    sharp corners,
    boxrule=2pt,               % 边框稍微加粗一点以匹配截图的视觉效果
    colframe=blue,             % 边框颜色改为纯蓝色
    colback=white,
    colbacktitle=white,
    coltitle=black,
    fonttitle=\bfseries,
    title={Low-Resource/Downscaling Tip: #1}, % 固定前缀，#1 接收自定义标题
    titlerule=0pt,
    overlay={
        % 调整了 xshift=10pt，使其与下方正文的 left=10pt 和 right=10pt 完美对齐
        \draw[black, line width=0.4pt] ([xshift=10pt, yshift=2pt]title.south west) -- ([xshift=-10pt, yshift=2pt]title.south east);
    },
    toptitle=8pt,
    bottomtitle=6pt,
    top=8pt,
    left=10pt,
    right=10pt,
    bottom=10pt
}

\begin{document}

\section{Assignment Overview}

\paragraph{What you will implement}

\begin{enumerate}[itemsep=0pt, parsep=0pt, topsep=0pt]
    \item Byte-pair encoding (BPE) tokenizer \hfill \tcbox[on line, boxsep=0pt, left=2pt, right=2pt, top=2pt, bottom=2pt, colback=green!10, colframe=green!50!black, fontupper=\tiny\bfseries]{DONE}
    \item Transformer language model (LM)
    \item The cross-entropy loss function and the AdamW optimizer
    \item The training loop, with support for serializing and loading model and optimizer state
\end{enumerate}

\paragraph{What you will run}

\begin{enumerate}[itemsep=0pt, parsep=0pt, topsep=0pt]
    \item Train a BPE tokenizer on the TinyStories dataset. \hfill \tcbox[on line, boxsep=0pt, left=2pt, right=2pt, top=2pt, bottom=2pt, colback=green!10, colframe=green!50!black, fontupper=\tiny\bfseries]{DONE}
    \item Run your trained tokenizer on the dataset to convert it into a sequence of integer IDs.
    \item Train a Transformer LM on the TinyStories dataset.
    \item Generate samples and evaluate perplexity using the trained Transformer LM.
    \item Train models on OpenWebText and submit your attained perplexities to a leaderboard.
\end{enumerate}

\newpage

\section{Byte-Pair Encoding (BPE) Tokenizer}
\subsection{The Unicode Standard}

In Python, you can use the \mintinline{python}{ord()} function to convert a single Unicode character into its integer representation.
The \mintinline{python}{chr()} function converts an integer Unicode code point into a string with the corresponding character.

\begin{problembox}{unicode1}{Understanding Unicode (1 point)}

    \begin{enumerate}[label=(\alph*), itemsep=0.5em, topsep=0pt]

        \item What Unicode character does \texttt{chr(0)} return? \\
              \textbf{Answer:} \mintinline{python}{'\x00'}

        \item How does this character's string representation \mintinline{python}{__repr__()} differ from its printed representation? \\
              \textbf{Answer:} The \mintinline{python}{__repr__()} is designed to be \textbf{unambiguous} and is mainly for developers, showing the technical details of the object. On the other hand, the printed representation is meant to be \textbf{readable} and user-friendly.

        \item What happens when this character occurs in text? It may be helpful to play around with the following in your Python interpreter and see if it matches your expectations:

              \textbf{Answer:} Code below.

              \begin{minted}[xleftmargin=1.5em, framesep=5pt]{pycon}
>>> chr(0)
'\x00'
>>> print(chr(0))

>>> "this is a test" + chr(0) + "string"
'this is a test\x00string'
>>> print("this is a test" + chr(0) + "string")
this is a teststring
\end{minted}

    \end{enumerate}

\end{problembox}

\newpage
\subsection{Unicode Encodings}

\begin{itemize}[itemsep=0pt, parsep=0pt, topsep=0pt]
    \item UTF-8: The Efficient Variable-Length Encoder, It uses 1 to 4 bytes per character.
    \item UTF-16: The Middle Ground, It uses either 2 or 4 bytes per character.
    \item UTF-32: The Simple Fixed-Length Encoder, It uses exactly 4 bytes for every character, regardless of what it is.
\end{itemize}

\begin{problembox}{unicode2}{Unicode Encodings (3 point)}

    \begin{enumerate}[label=(\alph*), itemsep=0.5em, topsep=0pt]

        \item What are some reasons to prefer training our tokenizer on UTF-8 encoded bytes, rather than UTF-16 or UTF-32? It may be helpful to compare the output of these encodings for various input strings.\\
              \textbf{Answer:} \mintinline{python}{'\x00'}

        \item Consider the following (incorrect) function, which is intended to decode a UTF-8 byte string into a Unicode string. Why is this function incorrect? Provide an example of an input byte string that yields incorrect results. \\
              \vspace{-1.5em}
              \begin{minted}[xleftmargin=1.5em, framesep=5pt]{python}
def decode_utf8_bytes_to_str_wrong(bytestring: bytes): 
    return "".join([bytes([b]).decode("utf-8") for b in bytestring])
\end{minted}

              \vspace{-1.5em}

              \begin{minted}[xleftmargin=1.5em, framesep=5pt]{pycon}
>>> decode_utf8_bytes_to_str_wrong("hello".encode("utf-8"))
'hello'
\end{minted}

              \textbf{Example:} \mintinline{python}{b'\xe4\xb8\xad'}.\\
              \textbf{Answer:} This function is incorrect because it attempts to decode each byte individually, which fails for any character encoded as a multi-byte sequence, resulting in a \mintinline{python}{UnicodeDecodeError}.

        \item Give a two byte sequence that does not decode to any Unicode character(s).\\
              \textbf{Example:} \mintinline{python}{b'\xff\xff'}.\\
              \textbf{Answer:} This sequence is invalid because the byte \mintinline{python}{0xFF} is explicitly forbidden in the UTF-8 standard and does not correspond to any valid lead or continuation byte in the encoding scheme.

    \end{enumerate}

\end{problembox}

\subsection{Subword Tokenization}

Sennrich et al. [2016] propose to use byte-pair encoding (BPE; Gage, 1994), a compression algorithm that iteratively replaces (``merges'') the most frequent pair of bytes with a single, new unused index.

\subsection{BPE Tokenizer Training}

\paragraph{Vocabulary initialization}

\paragraph{Pre-tokenization}
Use regular expressions to split the sentence into independent punctuation marks and independent items that may include spaces, and improve the merging effect by counting the occurrence frequency of these independent items.
\begin{minted}[xleftmargin=1.5em, framesep=5pt]{pycon}
>>> PAT = r"""'(?:[sdmt]|ll|ve|re)| ?\p{L}+| ?\p{N}+| ?[^\s\p{L}\p{N}]+|\s+(?!\S)|\s+"""
>>> # requires `regex` package
>>> import regex as re
>>> re.findall(PAT, "some text that i'll pre-tokenize")
['some', ' text', ' that', ' i', "'ll", ' pre', '-', 'tokenize']
\end{minted}

\paragraph{Compute BPE merges} During the BPE merging process, two \textbf{principles} must be followed:

\begin{itemize}[itemsep=0pt, parsep=0pt, topsep=0pt]
    \item Do not \textbf{cross} pre-tokenization boundaries (this is already guaranteed algorithmically);
    \item Handle issues with a \textbf{reproducible conflict resolution strategy}, such as selecting the lexicographically largest word pair to merge when the occurrence counts are the same.
\end{itemize}

\paragraph{Special tokens}

\subsection{Experimenting with BPE Tokenizer Training}

\paragraph{Parallelizing pre-tokenization} Using the provided segmentation code and the \mintinline{python}{multiprocessing} library, parallelize the tokenization statistics.

\paragraph{Removing special tokens before pre-tokenization}

\paragraph{Optimizing the merging step} After performing a statistical pass to obtain a tokenization table with frequencies, perform merging only within that tokenization table.

\begin{tipbox}{Profiling}
    You should use profiling tools like \texttt{cProfile} or \texttt{scalene} to identify the bottlenecks in your implementation, and focus on optimizing those.
\end{tipbox}

\end{document}